Unifying intra- and inter-specific variation in tropical tree mortality

### 2017-12-03 ms/manuscript.Rnw
Updated rstan citations

Eqs. \ref{eq:mortality_model}-\ref{eq:model} were fit to the data using Bayesian inference and with covariates for growth rate in previous census and wood density, as well as random effects. Growth rates were estimated from field measurements of diameter, which inevitably include observation error. In our dataset, \Sexpr{prop_neg_growth(BCI_model_dataset_true_dbh)}\% of estimated growth rates were negative. To ensure our mortality model was not biased by these unlikely values we first applied a probabilistic model to estimate \textit{``true growth''}, taking into account measurement error and the distribution of growth rate across the community (see Supplementary Material S1 for details; Fig. S1). The parameters $\alpha_s$, $\beta_s$ and $\gamma_s$ were modelled as a function of both wood density (measured at species level) and a species-level random effect:
\begin{equation}\label{eq:submodels}
\alpha_s = \alpha_{0,s} \, \left(\frac{\rho}{\rho_c}\right)^{\alpha_{1}},
\end{equation}
with similar formulations for $\beta_s$ and $\gamma_s$. Here $\alpha_{1}$ captures the effect of wood density $\rho$ on $\alpha_s$, while $\alpha_{0,s}$ captures any other species-level residual error not explained by wood density for species $s$, modelled as random realisations from log-normal distributions. The form of eq. \ref{eq:submodels} ensures that parameters remain positive; and on a log scale this equates to an additive linear model centered around $\rho_c$. We also centered growth rate $X_i$ at the lower 5\% quantile for both diameter increment and area growth (0.172 and 0.338, respectively), meaning $\alpha_s$ should be interpreted as the hazard rate when growth rate was very low. Weak priors on all hyper-parameters were set (see Supplementary Material S2 for details). Models were fit in R \Sexpr{paste(version$major, version$minor, sep='.')} using the package \texttt{rstan} \Sexpr{packageVersion("rstan")}\cite{StanDevelopmentTeam:2015uz} and employing some numerical optimisations (see Supplementary Material S3-S4 for details and example code). We executed three independent chains and in all cases modelled parameters converged within 2000 iterations. Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic \cite{Brooks:1998ju}. After discarding the first 2000 iterations as ‘burn in’, a further 2000 iterations were taken from the joint posterior. Species parameter estimates from the final model are shown in Figs. S3\textendash S5.

>>>>

Eqs. \ref{eq:mortality_model}-\ref{eq:model} were fit to the data using Bayesian inference and with covariates for growth rate in previous census and wood density, as well as random effects. Growth rates were estimated from field measurements of diameter, which inevitably include observation error. In our dataset, \Sexpr{prop_neg_growth(BCI_model_dataset_true_dbh)}\% of estimated growth rates were negative. To ensure our mortality model was not biased by these unlikely values we first applied a probabilistic model to estimate \textit{``true growth''}, taking into account measurement error and the distribution of growth rate across the community (see Supplementary Material S1 for details; Fig. S1). The parameters $\alpha_s$, $\beta_s$ and $\gamma_s$ were modelled as a function of both wood density (measured at species level) and a species-level random effect:
\begin{equation}\label{eq:submodels}
\alpha_s = \alpha_{0,s} \, \left(\frac{\rho}{\rho_c}\right)^{\alpha_{1}},
\end{equation}
with similar formulations for $\beta_s$ and $\gamma_s$. Here $\alpha_{1}$ captures the effect of wood density $\rho$ on $\alpha_s$, while $\alpha_{0,s}$ captures any other species-level residual error not explained by wood density for species $s$, modelled as random realisations from log-normal distributions. The form of eq. \ref{eq:submodels} ensures that parameters remain positive; and on a log scale this equates to an additive linear model centered around $\rho_c$. We also centered growth rate $X_i$ at the lower 5\% quantile for both diameter increment and area growth (0.172 and 0.338, respectively), meaning $\alpha_s$ should be interpreted as the hazard rate when growth rate was very low. Weak priors on all hyper-parameters were set (see Supplementary Material S2 for details). Models were fit in R \Sexpr{paste(version$major, version$minor, sep='.')} using the package \texttt{rstan} \Sexpr{packageVersion("rstan")}\cite{rstan2017} and employing some numerical optimisations (see Supplementary Material S3-S4 for details and example code). We executed three independent chains and in all cases modelled parameters converged within 2000 iterations. Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic \cite{Brooks:1998ju}. After discarding the first 2000 iterations as ‘burn in’, a further 2000 iterations were taken from the joint posterior. Species parameter estimates from the final model are shown in Figs. S3\textendash S5.

### 2017-12-03 ms/supplementary.Rnw
updated file names
Updated rstan citations

\noindent In Stan, the function $ \textrm{log1m\_exp}$ is defined so that $ \textrm{log1m\_exp}(x)$ has the same value as $\log(1.0-\exp(x))$, but the computation is faster, more accurate, and more stable \cite{StanDevelopmentTeam:2015uz}.

@@

\clearpage
\begin{figure*}[!hb]
\centering
\includegraphics[width=0.6\textheight]{figures/supp_fig1.png}
\caption{Observed vs predicted individual growth rates and stem diameters}
\label{figS1}
\end{figure*}

>>>>

\noindent In Stan, the function $ \textrm{log1m\_exp}$ is defined so that $ \textrm{log1m\_exp}(x)$ has the same value as $\log(1.0-\exp(x))$, but the computation is faster, more accurate, and more stable \cite{StanDevelopmentTeam:2017uz}.

@@

\clearpage

<<param-table,results="asis",center=TRUE, echo= FALSE>>= 
hyperparam_table(final_model)
@

\clearpage

\begin{figure*}[!hb]
\centering
\includegraphics[width=0.6\textheight]{figures/supp_fig1.png}
\caption{Observed vs predicted individual growth rates and stem diameters}
\label{figS1}
\end{figure*}
